---
title: "DATA2002 Group Project"
author: "T11A_Ontime_6"
date: "USYD | DATA2002 | November 2020"
output: 
  rmdformats::material:
    code_folding: hide
    fig_caption: yes
    number_sections: no
    self_contained: yes
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
---

````{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Packages Require to Run
#install.packages("tidyverse")
#install.packages("moonBook")
#install.packages("webr")
#install.packages("sjPlot")
#install.packages("GGally")
#install.packages("qtlcharts")
#install.packages("ggfortify")
#install.packages("caret")
```

# Page 1

## Abstract

--

## Introduction

--

## Data Set

--

# Page 2

## Analysis

```{r, warning=FALSE, message=FALSE}
#Load Required Packages
library(tidyverse)
library(moonBook)
library(webr)
library(sjPlot)
library(GGally)
library(qtlcharts)
library(ggfortify)
library(caret)

df = readxl::read_excel("data/Concrete_Data.xls") %>% janitor::clean_names()
```

```{r}
#Clean Column Names Manually
colnames(df)[1] = "cement"
colnames(df)[2] = "slag"
colnames(df)[3] = "flyash"
colnames(df)[4] = "water"
colnames(df)[5] = "plasticizer"
colnames(df)[6] = "coarseagg"
colnames(df)[7] = "fineagg"
colnames(df)[8] = "age"
colnames(df)[9] = "strength"
```

We will only use those data points with 28 age_days.

```{r}
dfx = df %>% filter(age == 28)

dfx
```

```{r}
#Checking Linearity
lm1 = lm(strength ~ cement, data = dfx)
lm2 = lm(strength ~ slag, data = dfx)
lm3 = lm(strength ~ flyash, data = dfx)
lm4 = lm(strength ~ water, data = dfx)
lm5 = lm(strength ~ plasticizer, data = dfx)
lm6 = lm(strength ~ coarseagg, data = dfx)
lm7 = lm(strength ~ fineagg, data = dfx)
#lm8 = lm(strength ~ age, data = df)

#Create a working dataframe
df1 = dfx

df1 = df1 %>% mutate(resid1 = lm1$residuals, fitted1 = lm1$fitted.values,
                     resid2 = lm2$residuals, fitted2 = lm2$fitted.values,
                     resid3 = lm3$residuals, fitted3 = lm3$fitted.values,
                     resid4 = lm4$residuals, fitted4 = lm4$fitted.values,
                     resid5 = lm5$residuals, fitted5 = lm5$fitted.values,
                     resid6 = lm6$residuals, fitted6 = lm6$fitted.values,
                     resid7 = lm7$residuals, fitted7 = lm7$fitted.values)

#Cement
p1 = ggplot(df1, aes(x = cement, y = strength)) + geom_point() +
     labs(x = "Cement (kg/m^3)", y = "Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)+geom_smooth(method = "loess", se = FALSE,col='red')
p1r = ggplot(df1, aes(x = cement, y = resid1)) + geom_point() +
      labs(x = "Cement (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Slag
p2 = ggplot(df1, aes(x = slag, y = strength)) + geom_point() +
     labs(x = "Slag (kg/m^3)", y = "Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)+geom_smooth(method = "loess", se = FALSE,col='red')
p2r = ggplot(df1, aes(x = slag, y = resid2)) + geom_point() +
      labs(x = "Slag (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#FlyAsh
p3 = ggplot(df1, aes(x = flyash, y = strength)) + geom_point() +
     labs(x = "FlyAsh (kg/m^3)", y = "Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)+geom_smooth(method = "loess", se = FALSE,col='red')
p3r = ggplot(df1, aes(x = flyash, y = resid3)) + geom_point() +
      labs(x = "FlyAsh (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Water
p4 = ggplot(df1, aes(x = water, y = strength)) + geom_point() +
     labs(x = "Water (kg/m^3)", y = "Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)+geom_smooth(method = "loess", se = FALSE,col='red')
p4r = ggplot(df1, aes(x = water, y = resid4)) + geom_point() +
      labs(x = "Water (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Plasticizer
p5 = ggplot(df1, aes(x = plasticizer, y = strength)) + geom_point() +
     labs(x = "Plasticizer (kg/m^3)", y = "Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)+geom_smooth(method = "loess", se = FALSE,col='red')
p5r = ggplot(df1, aes(x = plasticizer, y = resid5)) + geom_point() +
      labs(x = "Plasticizer (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Coarse Aggregate
p6 = ggplot(df1, aes(x = coarseagg, y = strength)) + geom_point() +
     labs(x = "Coarse Aggregate (kg/m^3)", y = "Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)+geom_smooth(method = "loess", se = FALSE,col='red')
p6r = ggplot(df1, aes(x = coarseagg, y = resid6)) + geom_point() +
      labs(x = "Coarse Aggregate (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Fine Aggregate
p7 = ggplot(df1, aes(x = fineagg, y = strength)) + geom_point() +
     labs(x = "Fine Aggregate (kg/m^3)", y = "Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)+geom_smooth(method = "loess", se = FALSE,col='red')
p7r = ggplot(df1, aes(x = fineagg, y = resid7)) + geom_point() +
      labs(x = "Fine Aggregate (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Age
#p8 = ggplot(df1, aes(x = age, y = strength)) + geom_point() +
#     labs(x = "Age (Days)", y = "Concrete Strength (mpa)") +
#     geom_smooth(method = "lm", se = FALSE)+geom_smooth(method = "loess", se = FALSE,col='red')
#p8r = ggplot(df1, aes(x = age, y = resid8)) + geom_point() +
#      labs(x = "Age (Days)", y = "Residual") +
#      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)

```

```{r}
#Plots
p1
#p1r
p2
#p2r
p3
#p3r
p4
#p4r
p5
#p5r
p6
#p6r
p7
#p7r
#p8
#p8r
```

```{r, warnings=FALSE}
#Residuals vs Fitted Values & QQ Plots
autoplot(lm1, which = 1:2)
autoplot(lm2, which = 1:2)
autoplot(lm3, which = 1:2)
autoplot(lm4, which = 1:2)
autoplot(lm5, which = 1:2)
autoplot(lm6, which = 1:2)
autoplot(lm7, which = 1:2)
#autoplot(lm8, which = 1:2)

```


```{r}
#tab_model(lm1, show.ci = FALSE) # p-value <0.001
#tab_model(lm2, show.ci = FALSE) # p-value <0.001
#tab_model(lm3, show.ci = FALSE) # p-value =0.001
#tab_model(lm4, show.ci = FALSE) # p-value <0.001
#tab_model(lm5, show.ci = FALSE) # p-value <0.001
#tab_model(lm6, show.ci = FALSE) # p-value <0.001
#tab_model(lm7, show.ci = FALSE) # p-value <0.001
#tab_model(lm8, show.ci = FALSE) # p-value <0.001
```
There is a significant relationship between each variable and concrete strength. 

Linearity - Slag, FlyAsh and Water fails the linearity check.

Homoscedasticity - For all the plots except Plasticizer, the spread looks reasonably constant over the range of values.

Normality - Based on the QQPlot, the variables follow a normal distribution. We could appeal to the Central Limit Theorem because we have a reasonable amount of observations meaning our inferences are approximately valid.

```{r}
#Log Transformations (Log - Linear)
df1 = df1 %>% mutate(lstrength = log(strength))

loglm1 = lm(lstrength ~ cement, data = df1)
loglm2 = lm(lstrength ~ slag, data = df1)
loglm3 = lm(lstrength ~ flyash, data = df1)
loglm4 = lm(lstrength ~ water, data = df1)
loglm5 = lm(lstrength ~ plasticizer, data = df1)
loglm6 = lm(lstrength ~ coarseagg, data = df1)
loglm7 = lm(lstrength ~ fineagg, data = df1)
#loglm8 = lm(lstrength ~ age, data = df1)

df1 = df1 %>% mutate(lresid1 = loglm1$residuals, lfitted1 = loglm1$fitted.values,
                     lresid2 = loglm2$residuals, lfitted2 = loglm2$fitted.values,
                     lresid3 = loglm3$residuals, lfitted3 = loglm3$fitted.values,
                     lresid4 = loglm4$residuals, lfitted4 = loglm4$fitted.values,
                     lresid5 = loglm5$residuals, lfitted5 = loglm5$fitted.values,
                     lresid6 = loglm6$residuals, lfitted6 = loglm6$fitted.values,
                     lresid7 = loglm7$residuals, lfitted7 = loglm7$fitted.values)
                     #lresid8 = loglm8$residuals, lfitted8 = loglm8$fitted.values)

#Cement
p1 = ggplot(df1, aes(x = cement, y = lstrength)) + geom_point() +
     labs(x = "Cement (kg/m^3)", y = "Log Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)
p1r = ggplot(df1, aes(x = cement, y = lresid1)) + geom_point() +
      labs(x = "Cement (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Slag
p2 = ggplot(df1, aes(x = slag, y = lstrength)) + geom_point() +
     labs(x = "Slag (kg/m^3)", y = "Log Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)
p2r = ggplot(df1, aes(x = slag, y = lresid2)) + geom_point() +
      labs(x = "Slag (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#FlyAsh
p3 = ggplot(df1, aes(x = flyash, y = lstrength)) + geom_point() +
     labs(x = "FlyAsh (kg/m^3)", y = "Log Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)
p3r = ggplot(df1, aes(x = flyash, y = lresid3)) + geom_point() +
      labs(x = "FlyAsh (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Water
p4 = ggplot(df1, aes(x = water, y = lstrength)) + geom_point() +
     labs(x = "Water (kg/m^3)", y = "Log Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)
p4r = ggplot(df1, aes(x = water, y = lresid4)) + geom_point() +
      labs(x = "Water (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Plasticizer
p5 = ggplot(df1, aes(x = plasticizer, y = lstrength)) + geom_point() +
     labs(x = "Plasticizer (kg/m^3)", y = "Log Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)
p5r = ggplot(df1, aes(x = plasticizer, y = lresid5)) + geom_point() +
      labs(x = "Plasticizer (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Coarse Aggregate
p6 = ggplot(df1, aes(x = coarseagg, y = lstrength)) + geom_point() +
     labs(x = "Coarse Aggregate (kg/m^3)", y = "Log Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)
p6r = ggplot(df1, aes(x = coarseagg, y = lresid6)) + geom_point() +
      labs(x = "Coarse Aggregate (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Fine Aggregate
p7 = ggplot(df1, aes(x = fineagg, y = lstrength)) + geom_point() +
     labs(x = "Fine Aggregate (kg/m^3)", y = "Log Concrete Strength (mpa)") +
     geom_smooth(method = "lm", se = FALSE)
p7r = ggplot(df1, aes(x = fineagg, y = lresid7)) + geom_point() +
      labs(x = "Fine Aggregate (kg/m^3)", y = "Residual") +
      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
#Age
#p8 = ggplot(df1, aes(x = age, y = lstrength)) + geom_point() +
#     labs(x = "Age (Days)", y = "Log Concrete Strength (mpa)") +
#     geom_smooth(method = "lm", se = FALSE)
#p8r = ggplot(df1, aes(x = age, y = lresid8)) + geom_point() +
#      labs(x = "Age (Days)", y = "Residual") +
#      geom_hline(yintercept = 0) + geom_smooth(method = "loess", se = FALSE)
```

```{r}
#Log-Linear Plots
p1
p1r
p2
p2r
p3
p3r
p4
p4r
p5
p5r
p6
p6r
p7
p7r
#p8
#p8r
```

Overall, a Log-Linear transformation looks slightly better than the Linear-Linear.


```{r}
#Remove Age Column
df = select(df, -age)
df2 = df

df2 = df2 %>% mutate(lstrength = log(strength))
df2 = select(df2, -strength)

M0 = lm(strength ~ 1, data = df) #Linear-Linear Null Model
M1 = lm(strength ~ ., data = df) #Linear-Linear Full Model
lM0 = lm(lstrength ~ 1, data = df2) #Log-Linear Null Model
lM1 = lm(lstrength ~ ., data = df2) #Log-Linear Full Model

#Linear-Linear Backward
step.back.aic = step(M1, direction = "backward", trace = FALSE)
round(summary(step.back.aic)$coef, 3)

#Linear-Linear Forward
step.fwd.aic = step(M0, scope = list(lower = M0, upper = M1), direction = "forward", trace = FALSE)
round(summary(step.fwd.aic)$coef, 3)

#Log-Linear Backward
log.step.back.aic = step(lM1, direction = "backward", trace = FALSE)
round(summary(log.step.back.aic)$coef, 3)

#Log-Linear Forward
log.step.fwd.aic = step(lM0, scope = list(lower = lM0, upper = lM1), direction = "forward", trace = FALSE)
round(summary(log.step.fwd.aic)$coef, 3)
```

```{r}
qtlcharts::iplotCorr(df)
```
```{r}
qtlcharts::iplotCorr(df2)
```



```{r}
#Linear-Linear Models
sjPlot::tab_model(
  step.fwd.aic, step.back.aic,
  show.ci = FALSE, 
  show.aic = TRUE,
  dv.labels = c("Linear-Linear Forward Model",
                "Linear-Linear Backward Model")
)
```

```{r}
autoplot(step.fwd.aic, which = 1:2)
```
```{r}
cv_linear= train(
  strength ~ cement+plasticizer+slag+water+flyash,df,
  method='lm',
  trControl=trainControl(method='cv',number=10,verboseIter=FALSE)
)
cv_linear
```


```{r}
#Log-Linear Models
sjPlot::tab_model(
  log.step.fwd.aic, log.step.back.aic,
  show.ci = FALSE, 
  show.aic = TRUE,
  dv.labels = c("Log-Linear Forward Model",
                "Log-Linear Backward Model")
)
```

The Log-Linear forward and backward models starting with 7 variables, produced a R^2 adjusted value of 0.383 and AIC of 1207.299 and 1207.138 respectively.

```{r}
autoplot(log.step.fwd.aic, which = 1:2)
autoplot(log.step.back.aic, which = 1:2)
```


```{r}
cv_log= train(
  lstrength ~ cement+water+coarseagg+fineagg+plasticizer,df2,
  method='lm',
  trControl=trainControl(method='cv',number=10,verboseIter=FALSE)
)
cv_log
```

```{r}

```


```{r}
#Stepwise Variable Selection
lM2 = lm(lstrength ~ ., data = df2)
drop1(lM2, test = "F")
lM3 = update(lM2, . ~ . - coarseagg)
drop1(lM3, test = "F")
lM4 = update(lM3, . ~ . -fineagg)
drop1(lM4, test = "F")

sjPlot::tab_model(
  lM4,
  show.ci = FALSE, 
  show.aic = TRUE,
  dv.labels = c("Stepwise Variable Selection")
)
```

The stepwise variable selection produced a R^2 adjusted value of 0.382 and AIC of 1208.186. The stepwise is a better model as it has one lesser variable and the AIC differs only by 1.048 indicating somewhat equally fitting.


--

## Results

--

## Discussion and Conclusion

--

# References & Additional Figures

